`mrpkit`: A reproducible, opinionated, workflow for multilevel regression with post-stratification in R

### 2022-01-24 paper/paper.Rmd
Something is better than nothing

thanks: "**Feedback:** Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: lauren.kennedy729@gmail.com. **Funding statement:** We have no direct financial disclosures. **Conflicts statement:** We have no conflicts to disclose. **Code and data accessibility statement:** Code and data are available at: https://github.com/lauken13/mrpkit. We are unable to share direct access to the datasets used in the 2020 US Presidential example in this paper, however they are freely available and we provide detailed instructions for how to obtain them. **Author contributions:** Kennedy had the original idea. Gabry, Kennedy, and Morris made architectural and other high-level decisions. Gabry and Kennedy wrote most of the code and tests in the package. Amaliah wrote the vignette and contributed tests. Alexander wrote the first draft of this paper. All authors contributed to writing the paper, and approved the final version."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: |
    The R package `mrpkit` implements a structured workflow for Multilevel Regression and Poststratification (MRP). The package helps appropriately set-up the survey and post-stratification datasets and then establish a relationship between different variables in these datasets. A substantial amount of the work to do this is specified by the methods in the package, saving time, enhancing reproducibility, and reducing the risk of coding errors. The package natively supports multilevel binomial and Bernoulli models fit with `lme4` and `Stan` (via `brms` and `rstanarm`) and allows for the use of custom modeling functions. Summaries and visualizations of the resulting post-stratified estimates can be conducted using the methods in the package. This implementation provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching. Using `mrpkit` proactively addresses many common issues and makes it possible to conduct MRP in an efficient way.

@@

The process of conducting MRP involves considering some survey outcome as being due to a combination of many different cells that are typically demographic or geographic in nature, estimating this relationship using a multilevel regression model and then aggregating cell-level estimates to match the population [@wang2015forecasting]. The survey of interest would typically be a regular survey, such as a political poll of, say, 1,000 respondents, but it could also be a larger survey, such as the American National Election Studies, or similar. The post-stratification dataset would typically be a larger survey, such as, in the case of the US, the American Community Surveys, or a census. The correspondences need to be established for each possible answer for each question, for instance, the age-groups used in both surveys need to be homogenized. The model would typically be a Bayesian hierarchical model, but in principle any variety could be used.

@@

Our package complements existing packages such as `survey` [@lumleyone; @lumleytwo; @lumleythree], and `DeclareDesign` [@citeDeclareDesign]. The `survey` package is well-established and allows a user to specify a survey design, such as clustered, stratified, etc, generate summary statistics, and conduct sub-population analysis. The `srvyr` package [@citesrvyr] is a variant of `survey` that adds a syntax similar to `dplyr` [@citedplyr]. The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design aspects including: data generation, creating treatment and control groups, and sampling. In general, these packages are focused on designing, implementing, and simulating from, surveys. Ours is focused on what is needed for MRP, and we draw on their packages where possible.

@@

The main alternative approach is for MRP to be conducted on a case-specific basis. This means that a researcher who is interested in MRP estimates writes the code needed to obtain, clean, prepare, analyze, and ultimately interpret the estimates. Again, while there is nothing inherently wrong with this approach, successfully conducting MRP requires dealing with many small issues. Each of these is small but getting them wrong can have large effects that are potentially unnoticed on the estimates. It is also easy to overlook steps, or to not appropriately document them. `mrpkit` is prescriptive about the steps that must be taken, and ensures that all aspects are documented.

@@

- A professor wants to introduce MRP to a class that might not be especially experienced with R, nor have time to develop the necessary skills, instead wanting to focus on the estimates and applications.
- A researcher with domain or survey expertise, but who does not have the resources to ensure that the code has been appropriately tested.
- A researcher who is experienced with MRP needs to run many different models on the same dataset.

@@

Having installed and loaded the package, a user first needs to get their survey and post-stratification dataset into R. This means ensuring that it is in a rectangular format such as a CSV and then reading it in as a dataframe. The `mrpkit` package does not explicitly contain methods for this, instead a user could use `readr::read_csv()` [@citereadr] or `haven::read_dta()` [@citehaven].

A user will need two datasets. At this point, a user should be keenly aware of the class of the variables as this becomes important later in the workflow. This is because the `mrpkit` package will convert all variables to factor variables.

@@

1. `data`: The name of the dataset, i.e., the survey or post-stratification data that will be transformed to a `SurveyData` object. By default, this method uses all of the factor, character, and binary variables in the dataset to create questions and responses, unless a user specifies the list of questions and responses that are of interest.
2. `questions`: The actual questions asked in the survey form or questionnaire for each of the columns in the dataset For instance, a column called 'age' may contain the results of a question 'Please identify your age-group'.
3. `responses`: All possible responses for each of the questions. For instance, possible responses in the 'age' column could be: '18-25', '26-35', '36-45', '46-55', '56-65', '66-75', and '76-90'.
4. (Optional) `weights`: If the weight is specified as a column in the dataset, then this argument should be specified as a string.
5. (Optional) `design`: If there is a design formula for the survey, then this can be specified using the notation of the `survey` package [@lumleyone; @lumleytwo; @lumleythree].

@@

To initialize a new `SurveyMap` object a user can use the `SurveyMap$new` method. This takes `SurveyData` and `QuestionMap` objects as argument.

```{r map-object}
# Create a new SurveyMap object
ex_map <- SurveyMap$new(sample = box_pref, population = poststrat_obj,
                        q_age, q_educ, q_gender, q_party_pref
                        )

@@

The next step is to generate the population estimate or group estimate using the `SurveyFit$aggregate` method. This method takes two arguments, namely the post-stratification estimate and the variable whose level to which the estimated value would be aggregated to. In this example, we want to aggregate the estimated value by the level of `age` and `state`. If the variable is not specified, then this method would generate a population estimate.

@@

# Implementation and technical notes {#implementation}

There are a variety of technical notes related to this package and the architectural decisions that were made.

Section \@ref(implementation) discusses some of the implementation issues and technical notes related to the decisions that were made

>>>>

thanks: "**Feedback:** Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: lauren.kennedy729@gmail.com. **Funding statement:** We have no direct financial disclosures. **Conflicts statement:** We have no conflicts to disclose. **Code and data accessibility statement:** Code and data are available at: https://github.com/lauken13/mrpkit. We are unable to share direct access to the datasets used in the 2020 US Presidential example in this paper, however they are freely available and we provide detailed instructions for how to obtain them. **Author contributions:** Kennedy had the original idea. Gabry, Kennedy, and Morris made architectural and other high-level decisions. Gabry and Kennedy wrote most of the code and tests in the package. Amaliah wrote the vignette and contributed tests. Alexander wrote the first draft of this paper. All authors contributed to writing the paper, and approved the final version."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: |
    The R package `mrpkit` implements a structured workflow for Multilevel Regression and Poststratification (MRP). The package helps appropriately set-up the survey and post-stratification datasets and then establish a relationship between different variables in these datasets. A substantial amount of the work to do this is specified by the package's methods, saving time, enhancing reproducibility, and reducing the risk of coding errors. The package natively supports multilevel binomial and Bernoulli models fit with `lme4` and `Stan` (via `brms` and `rstanarm`) and allows for the use of custom modeling functions. Summaries and visualizations of the resulting post-stratified estimates can be conducted using the package's methods. This implementation provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching. Using `mrpkit` proactively addresses many common issues and makes it possible to conduct MRP in an efficient way.

@@

```{r laurenflowchart, echo = FALSE, fig.cap = "Example of typical MRP workflow", out.width = '90%'}
knitr::include_graphics(here::here("paper/mrpkit_workflow.pdf"))
```

The process of conducting MRP involves considering some survey outcome as being due to a combination of many different cells that are typically demographic or geographic in nature, estimating this relationship using a multilevel regression model and then aggregating cell-level estimates to match the population Figure \@ref(fig:laurenflowchart). The survey of interest would typically be a regular survey, such as a political poll of, say, 1,000 respondents, but it could also be a larger survey, such as the American National Election Studies, or similar. The post-stratification dataset would typically be a larger survey, such as, in the case of the US, the American Community Surveys, or a census. The correspondences need to be established for each possible answer for each question, for instance, the age-groups used in both surveys need to be homogenized. The model would typically be a Bayesian hierarchical model, but in principle any variety could be used.

@@

Our package complements existing packages such as `survey` [@lumleyone; @lumleytwo; @lumleythree], and `DeclareDesign` [@citeDeclareDesign]. The `survey` package is well-established and allows a user to specify a survey design, such as clustered, stratified, etc, generate summary statistics, and conduct sub-population analysis. The `srvyr` package [@citesrvyr] is a variant of `survey` that adds a syntax similar to `dplyr` [@citedplyr]. The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design. These include: data generation, creating treatment and control groups, and sampling. In general, these packages are focused on designing, implementing, and simulating from, surveys. Ours is focused on what is needed for MRP, and we draw on their packages where possible.

@@

The main alternative to using `mrpkit` is for MRP to be conducted on a case-specific basis. This means that a researcher who is interested in MRP estimates writes the code needed to obtain, clean, prepare, analyze, and ultimately interpret the estimates. Again, while there is nothing inherently wrong with this approach, successfully conducting MRP requires dealing with a variety of issues. Each of these is small, but getting them wrong can have large effects on the estimates that are potentially unnoticed. It is also easy to overlook steps, or to not appropriately document them. `mrpkit` is prescriptive about the steps that must be taken when conducting MRP, and ensures that all aspects are documented.

@@

- A professor wants to introduce MRP to a class that might not be especially experienced with R, nor have time to develop the necessary skills, instead wanting to focus on the estimates and applications.
- A researcher with domain or survey expertise, but who does not have the resources to ensure that the code has been appropriately tested.
- A researcher who is experienced with MRP needs to run many different models on the same, or similar, dataset/s.

@@

Having installed and loaded the package, a user first needs to get their survey and post-stratification dataset into R. This means ensuring that it is in a rectangular format such as a CSV and then reading it in as a dataframe. The `mrpkit` package does not explicitly contain methods for this. Instead a user could use `readr::read_csv()` [@citereadr] or `haven::read_dta()` [@citehaven].

A user will need two datasets. At this point, a user should take note of the class of the variables as this becomes important later in the workflow. This is because the `mrpkit` package may convert some variables to factors.

@@

1. `data`: The name of the dataset, i.e., the survey or post-stratification data that will be transformed to a `SurveyData` object. By default, this method uses all of the factor, character, and binary variables in the dataset to create questions and responses, unless a user specifies the list of questions and responses that are of interest.
2. `questions`: The actual questions asked in the survey form or questionnaire for each of the columns in the dataset. For instance, a column called 'age' may contain the results of a question 'Please identify your age-group'.
3. `responses`: All possible responses for each of the questions. For instance, possible responses in the 'age' column could be: '18-25', '26-35', '36-45', '46-55', '56-65', '66-75', and '76-90'.
4. (Optional) `weights`: If the weight is specified as a column in the dataset, then this argument should be specified as a string.
5. (Optional) `design`: If there is a design formula for the survey, then this can be specified using the notation of the `survey` package [@lumleyone; @lumleytwo; @lumleythree].

@@

To initialize a new `SurveyMap` object a user can use the `SurveyMap$new` method. This takes `SurveyData` and `QuestionMap` objects as arguments.

```{r map-object}
# Create a new SurveyMap object
ex_map <- SurveyMap$new(sample = box_pref,
                        population = poststrat_obj,
                        q_age,
                        q_educ,
                        q_gender,
                        q_party_pref)

@@

The next step is to generate the population estimate or group estimate using the `SurveyFit$aggregate` method. This method takes two arguments, namely the post-stratification estimates and the variable whose level to which the estimated value would be aggregated to. In this example, we want to aggregate the estimated value by the level of `age` and `state`. If the variable is not specified, then this method would generate a population estimate.

@@
